---
title: "Homework-03"
author: "Cassidy Thorpe"
date: "2026-05-26"
format: pdf
execute:
  message: false
  warning: false
---

# Set Up

```{r}
#| label: packages

# load in packages
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(ggeffects)
```

```{r}
#| label: data

# store kelp data as an object called kelp
kelp <- read_csv(here("data","temp-kelp.csv"))
# store egg data as an object called egg_data
egg_data <- read_csv(here("data", "personal-data-project - Sheet1.csv"))
```

# Problems

## Problem 1: Giant Kelp Fronds

### a. An appropriate test

To determine the strength of the relationship between temperature and elongation rate, you could do a Pearson's correlation and a Spearman rank correlation. A Pearson’s correlation is parametric and is best for understanding the strength of the correlation between variables when
there is a linear relationship between continuous variables. A Spearman’s rank correlation is non-parametric and useful for understanding the strength of the correlation between variables that show a
monotomic relationship between variables.

### b. Create a visualization

```{r}
#| label: scatterplot

# base layer: ggplot
ggplot(data = kelp, # use kelp data frame
       aes(x = temp_c, # x-axis: temperature (C)
           y = kelp_elong)) + # y-axis: kelp elongation rate (cm/day)
  # 1st layer: show raw data points
  geom_point(color = "darkgreen", # points = dark green
             alpha = 0.6) + # points = slightly transparent
  # label and and y axes
  labs(x = "Ocean Temperature(C)",
       y = expression(paste("Predicted Elongation Rate (cm ", day^-1,")")),
       # add title 
       title = "Kelp elongation rate is faster in colder waters") + 
  # change theme
  theme_bw() +
  # remove legend
  theme(legend.position = "none")
```

### c. Check assumptions and run test

**Create Linear Model**

```{r}
#| label: model

# create new object called kelp model to store linear model
# function: lm
# formula: dependent ~ independent, data = data frame
kelp_model <- lm(kelp_elong ~ temp_c, data = kelp) 
```

```{r}
#| label: summary

# show summary of linear model
summary(kelp_model)
flextable::as_flextable(kelp_model)
```

**Check Assumptions**

```{r}
#| label: asumptions

# create 2x2 grid to show diagnostic plot
par(mfrow = c(2,2))
# show diagnostic plots
plot(kelp_model)
```

The QQ residuals plot shows the data points following the line. This suggests the distribution of the residuals is normal. 
The residuals vs. fitted and scale locations plots don't seem to show a pattern. The data seems to be randomly and evenly distributed  which indicates that the residuals are homoscedastic. 
The residuals vs leverage plot does not show points outside of the dashed lines which indcates that there are no outliers influencing the model. 

**Run Test**

```{r}

kelp_cor <- cor.test(kelp$temp_c, kelp$kelp_elong, method = "pearson")
show(kelp_cor)
```

### d. Results communication

To evaluate the strength of the relationship between temperature and kelp frond elongation, I used Pearson's correlation (r). This test is a good choice because both ocean temperature and elongation rate are continuous variables, and the diagnostic checks show that the data satisfies the assumptions of normality and linearity. I suspect a linear relationship between ocean temperature and kelp frond elongation rate

I found a strong negative  relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r = -0.7, t(30) = -5.189, p < 0.001, $\alpha$ = 0.05).







