---
title: "Homework-03"
author: "Cassidy Thorpe"
date: "2026-05-26"
format: pdf
---

# Set Up

```{r}
#| label: packages

# load in packages
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(ggeffects)
```

```{r}
#| label: data

# store kelp data as an object called kelp
kelp <- read_csv(here("data","temp-kelp.csv"))
# store egg data as an object called egg_data
egg_data <- read_csv(here("data", "personal-data-project - Sheet1.csv"))
```

# Problems

## Problem 1: Giant Kelp Fronds

### a. An appropriate test

To determine the strength of the relationship between temperature and elongation rate, you could do a Pearson's correlation and a Spearman rank correlation. A Pearson’s correlation is parametric and is best for understanding the strength of the correlation between variables when
there is a linear relationship between continuous variables. A Spearman’s rank correlation is non-parametric and useful for understanding the strength of the correlation between variables that show a
monotomic relationship between variables.

### b. Create a visualization

```{r}
#| label: scatterplot

ggplot(data = kelp,
       aes(x = temp_c,
           y = kelp_elong)) +
  geom_point(color = "darkgreen",
             alpha = 0.6) + 
  labs(x = "Ocean Temperature(C)",
       y = expression(paste("Predicted Elongation Rate (cm ", day^-1,")")),
       title = "Kelp elongation rate is faster in colder waters") + 
  theme_bw() +
  theme(legend.position = "none")
```

```{r}
#| label: scatterplot-plus

kelp_model <- lm(kelp_elong ~ temp_c, data = kelp) 
kelp_predict <- ggpredict(kelp_model, terms = "temp_c")
# base layer: ggplot
ggplot(data = kelp_predict, 
       aes(x = x, # x-axis: x
           y = predicted # y-axis: predicted 
)) +
  # 1st layer: geom_point to show raw data points
  geom_point(data = kelp, 
             aes(x = temp_c,
                 y = kelp_elong, color = "darkgreen", alpha = 0.5),
             color = "darkgreen", # point color = dark green
             alpha = 0.5) + # make points slightly transparent
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = "lightgreen",
              alpha = 0.3) + 
  geom_line(color = "darkgreen") +
  labs(x = "Ocean Temperature (C)",
       y = expression(paste("Predicted Elongation Rate (cm ", day^-1,")")),
       title = "Colder ocean temepratures predict faster elongation rate") + 
  theme_bw()
```

```{r}

```







